Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2019-12-09 paper/manuscript.tex
working my way slowly through citations

\paragraph{Several choices of restraints are possible.}
\todo[inline, color={blue!20}]{@author: Check references
In practice, a variety of types of restraints are common, from simple harmonic distance restraints between the ligand and the protein~\cite{mobley2006use}, to flat-bottom restraints which work similarly but only exert a force if the ligand leaves a specific region~\cite{}.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{paper/figures/fig10_equib_detection/fig10_equib.pdf}
    \caption{{\bf Automatic partitioning into equilibration and production regions.}
    \emph{Top:} The average (black line) standard deviation (shaded region) of the reduced potential $u^*$ over many independent replicate simulations started from the same initial conditions show a significant initial transient before relaxing to the true average potential energy.
    \emph{Middle:} A cumulative average (red) of the entire simulation data demonstrates a bias in the estimate of the average potential $<u>$ that persists far beyond the initial transient relaxation time, while averaging a production region that omits the initial transient (blue) eliminates this bias.
    \emph{Bottom:} A simple automated approach to determination of the equilibration boundary $t_0$ suggested in~\cite{chodera2016simple} examines the number of effectively uncorrelated samples $N_\mathrm{eff}_{[t_0,T]} = (T-t_0) / g_{[t_0,T]}$ within the proposed production region $[t_0,T]$ by computing the statistical inefficiency $g_{[t_0,T]}$ for that simulation interval.
    $N_\mathrm{eff}_{[t_0,T]}$ has the general behavior of increasing linearly as $t_0$ moves backwards from $T$ until the initial transient relaxation is encountered, at which point it plunges.
    The optimal equilibration boundary $t_0$ is selected to maximize the number of uncorrelated samples.
    \emph{Figure adapted from~\cite{chodera2016simple}.}
    \todo[inline]{JDC will update this figure with something similar for a real alchemical free energy calculation.}
    }
    \label{fig:automatic-equilibration-detection}
\end{figure}

@@

A simple approach to automatically partitioning simulation data into equilibration and production regions is described in~\cite{chodera2016simple} (illustrated in Figure~\ref{fig:automatic-equilibration-detection}).
Suppose we have a simulation of length $T$ consisting of correlated data.
Here, the goal of the post-processing step is to select the equilibration boundary $t_0 \in [0, T]$ so as to \emph{maximize} the number of effectively uncorrelated samples remaining in the production region $N_{[t_0,T]}$, which is defined as
\begin{eqnarray}
N_{[t_0,T]} &=& \frac{T - t_0}{g_{[t_0,T]}}
\end{eqnarray}
where $g_{[t_0,T]}$ is the \emph{statistical inefficiency} of a timeseries $a_t$, described in more detail below.
Conveniently, this procedure also produces the information necessary to decorrelate the simulation data for estimating the free energy differences, a requisite next step in analysis.
This approach is implemented within the MBAR~\cite{mbar} and alchemlyb~\cite{alchemlyb} packages, and is highly recommended for standard practice.

@@

Robust estimation of $C_t$ for $t \sim T$ is difficult due to growth in statistical error, so common estimators of $g$ make use of several additional properties of $C_t$ to provide useful estimates (see \emph{Practical Computation of Statistical Inefficiencies} in~\cite{chodera2016simple} for a detailed discussion).
\todo[inline]{JDC: Describe recommended statistical inefficiency computation algorithm.}
We recommend using the robust statistical inefficinecy computation routines available within the MBAR~\cite{mbar} and alchemlyb~\cite{alchemlyb} packages.

\paragraph{Subsampling data to generate uncorrelated samples}
Once the statistical inefficiency $g$ has been estimated, it is straightforward to subsample the correlated timeseries simulation data to produce effectively uncorrelated data that can ba fed to estimators of the free energy.
Suppose the correlated timeseries is $\{a_t\}_{t=1}^T$; we can form a new timeseries of $N_\mathrm{eff} \approx T / g$ effectively uncorrelated samples by selecting a subset of indices $\{ \: t = \mathrm{round}((n-1) \, g) \: | \: n \in \range 1,\ldots,N \: \}$ where $\mathrm{round(x)}$ denotes rounding to the nearest integer.

@@

\paragraph{Cautions and considerations}
Reliable estimation of the statistical inefficiency is difficult, and estimates will not generally be as precise (in a relative error sense) as averages.
To ensure there is sufficient data available for reliable decorrelation and estimation of free energy differences, it is recommended that the effective number of uncorrelated samples $N_\mathrm{eff} \geapprox 50$.

@@

The simplest method for calculating free energy
differences from simulations is the so-called \textit{Zwanzig
relationship}~\cite{orig.TPT}, also called one-sided exponential reweighting (EXP), or simply free energy perturbation, thought that term later grew to encompass all ways of calculating free energy differences.

The free energy between two states defined by two different potential energy functions 
$U_0(\vec{q})$ and $U_1(\vec{q})$ over coordinate space $\vec{q}$ can be calculated as:
\begin{eqnarray}
\Delta f_{01} & = & \ln \expect{e^{-(U_1(\vec{q}) - U_0(\vec{q}))}}_0 =  \ln \expect{e^{-\Delta U(\vec{q})} }_0
\end{eqnarray}\label{eqn.zwanzig}
In words, we take the samples generated during our run with the potential energy function $U_0(\vec{q})$, recalculate what the energy difference would be to switch to potential energy function $U_1(\vec{q})$, and exponentially average them to get the free energy.  Here, the exponential average consists in multiplying by -1, exponentiating, averaging, taking the logarithm, and multiplying by -1 again.  This almost requires no extra code functionality to perform; one need only to save a full precision trajectory, and run an unmodified molecular simulation code using the $U_1$ in order to calculate the new energies.  The analysis can be written in a line of code.  We note that this method is even more general, in that the instantaneous work to change the potential energy function from $U_0$ to $U_1$ can be replaced by the non-reversible work $W$ to make the same change some conditions~\cite{xx} 
Although the Zwanzig equation is formally correct (as long as the two states considered sample the same phase space volume), it has some very important numerical issues that mean that it is usually when performing a standard free energy calculation for a small molecule.~\cite{shirts.comparison,LuND:Appmcf}  One can show that if the standard deviation of the difference $\Delta U(\vec{q}) = U_1(vec{q})-U_2(\vec{q})$ over all sampled $\vec{q}$ is large (which in this case, means only several times $k_BT)$, then very few samples contribute to the average, and the answer will be both biased and extremely noisy~\cite{xx}.  There are a number of applications where it is useful, but those usually involve investigating the dependence of the free energy on small changes in parameters, not changing molecules or even single atoms in molecules~\cite{xx}

@@

Many studies have demonstrated both the
theoretical and practical superiority of BAR over EXP in molecular
simulations~\cite{shirts.comparison,LuND:Appmcf}, and BAR converges to EXP in the limit that all samples are from a
single state~\cite{bennettCH,shirts.bennett}. Significantly less
overlap between the configurational space of each state is required to
converge than for EXP, though some overlap must still exist.

@@

Various numerical integration schemes are possible, but the trapezoid
rule provides a simple and robust scheme.  All types of
numerical integration can be written as:
\[ \Delta f \approx \sum_{k=1}^{K} w_k
\expect{\frac{du(\lambda,\vec{q})}{d\lambda}}_{k} \] where the weights
$w_k$ correspond to a particular choice of numerical integration.
Researchers have tried a large number of different integration
schemes~\cite{resat.mezei.93,jorge_effect_2010,shyu_reducing_2009}. For However, many other integration choices require specific choices of $\lambda$
to minimize bias, which makes them unsuitable when the intermediates
have widely-varying levels of uncertainty. For example, integrating a cubic spline interpolation provided negligble benefits~\cite{paliwal.benchmark}. For starting researchers, we therefore recommend a simple trapezoid rule scheme, as it allows for maximal flexibility in which values of $\lambda$ are simulated.  Because fitting to higher
order polynomials can have numerical instabilities for some energy functions, and because alternate functional forms might only be appropriate
with some types of transformations, expertise and experience is
required to perform such numerical integration modifications.  In practice, adding 2-3 more intermediates is sufficient to match the performance of these more complicated numerical quadrature schemes.

@@

MBAR is provably the lowest variance asymptotically unbiases estimator of the free energies given the energies of the samples~\cite{zqtan}. It also provides an uncertainty estimate, which has been shown to be accurate as long as there are sufficient samples at each state~\cite{paliwal:benchmark}.

MBAR can also be thought of as the Zwanzig estimator of the free energy to state $i$ where the sampled distribution is the \textit{mixture distribution} of all the other samples thrown together in one "pot", defined as $p_m(\vec{q} = N^{-1} \sum_k  N_k \exp(f_k-u_k\vec{q})$, the weighted average of all the individual normalized probability distributions that are performed.~\cite{shirts.prl}.

@@

\begin{itemize}
\item MBAR recommended if all energy differences are available. It is provably the lowest variance free energy estimate given samples from multiple states.
\item BAR is essentially just as good as MBAR for highly optimized $\lambda$ intermediates.  Specifically, if the $\lambda$s are chosen such that intermediate states have moderate overlap with their neighbors (i.e. between $i$ and $i+1$ and between $i$ and $i-1$, they will \textit{not} have significant overlap with their next nearest neighbors $i+2$ and $i-2$. Thus MBAR doesn't actually get significant information from these energy differences, so one might as well not even calculate them, and just perform BAR between nearest neighbors.~\cite{paliwal.benchmark} 
\item TI usually gives similar values as MBAR implemented with sufficient numbers of intermediates, but quadrature error hard to estimate beforehand.~\cite{paliwal.benchmark}
\item WHAM is an approximation to MBAR, and there are no compelling reasons it should be used. 
\item Other variants useful in special circumstances (e.g. Z. Tan stochastic version), but beyond the scope of a Best Practices article.
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%
%              Terminology                  %
%%%%%%%%%%%%%%%%%%%%
\section{Terminology and abbreviations}
\label{sec:tem-abbrev}
\begin{itemize}
\item Feature Box covering major technical terms and abbreviations
\item Examples:
\begin{itemize}
\item EXP, BAR, MBAR
\item Double decoupling, single-topology, dual-topology, hybrid-topology, coupled-topology
\item FEP (free energy peturbation), alchemical, AFE (alchemical free energy)
\end{itemize}
\end{itemize}
%%%%%%%%%%%%%%%%%%%%
%                Software                    %
%%%%%%%%%%%%%%%%%%%%
\section{Available software -- a summary}
\todo[inline]{Samar: write section}
\label{sec:software}
There are several softwares availble for carrying out alchemical free energy calculations. In this section we discuss some of the popular commercial and free tools along with their features. 
\begin{itemize}
\item Commercial:
   \begin{itemize}
    \item FEP+ is a tool offered by Schrodinger Inc. under a commercial license. It has an intuitive GUI which makes it easier for non-experts to run alchemical free energy calculations and analyze the results. It runs DESMOND MD package under the hood and hence parallelizes very well on the GPUs. 
    \end{itemize}
\item Free or low-cost for academics / commercial for industry:
	\begin{itemize}
	\item CHARMM has a variety of tools developed over the years. PERT module can be used to define initial and final states and define the intermediate lambda points. FREN and BAR modules can be used to analyze the data after the MD run. Lambda-dynamics based free energy calulation can be carried out using the BLOCK module.  
	\item TIES and AMBER FEW? (Peter Coveney)
	\item AMBER, including its new pmemd.cuda version support free energy calculation. 
	\end{itemize}
\item Free (libre) open source:
	\begin{itemize}
	\item PLUMED is an open source tool which enables the usage of a variety of MD engines. It is designed as a plugin for MD packages such that it analyzes the trajectory on the fly. It also offers a VMD based plugin for the computation of collective variables.   	
	\item SIRE  is a free, open source, multiscale molecular simulation framework, written to allow computational modellers to quickly prototype and develop new algorithms for molecular simulation and molecular design. 
	\item YANK is a tool developed by John Chodera and group on the top of OpenMM MD package. It allows the users to write their inputs in easy-to-use YAML format.
	\item GROMACS is a molecular simulation package with a significant number of free energy methods implementations. The LiveCOMS GROMACS tutorial has an example free energy calculation~\cite{LiveCOMS-gromacs}.
	\item pmx for mutations
	\end{itemize}
\item Setup tools
	\begin{itemize}
	\item FESetup: AMBER, gromacs, Sire
	\item Lomap/Lomap2 : Relative alchemical transformation graph planning
	\item CHARMM-GUI is a web based tool for setting up a variety of MD simulations. It can be used to generate CHARMM sctipts for solvation and ligand-binding free energy calculations.
	\end{itemize}
\item Analysis tools:
	\begin{itemize}
	\item Free Energy Workflows: Sire-specific free energy map analysis using weighted path averages
	\url{https://github.com/michellab/freenrgworkflows}
	\item Alchemlyb: Multipackage free energy analysis
	\url{https://github.com/alchemistry/alchemlyb}
	\item pymbar: MBAR implementation, but have to roll your own analysis wrapper
	\url{https://github.com/choderalab/pymbar}
	\end{itemize}
\end{itemize}

@@

\begin{tabular}{cccc}
\caption{Overview of benchmark datasets}
\textbf{Publication} & \textbf{Targets} & \textbf{Ligands} & \textbf{Force Field} \\
\hline
\multicolumn{4}{|c|}{D3R Grand Challenges~\cite{D3R2019Dec}} \\
\hline
GC3~\cite{gaieb2019d3r} & 6 & 266 & various \\
GC2~\cite{gaieb2018d3r} & 1 & 102 & various \\
GC2015~\cite{gathiaka2016d3r} & 2 & 215 & various \\
\hline
\multicolumn{4}{|c|}{SAMPL Challenges~\cite{SAMPL}} \\
\hline
SAMPL6~\cite{rizzi2018overview} & 3 & 21 & various \\
SAMPL5~\cite{yin2017overview} & 3 & 22 & various \\
SAMPL4~\cite{muddana2014sampl4} & 2 & 23 & various \\
\hline
\multicolumn{4}{|c|}{Schrödinger Datasets} \\
\hline
FEP+ Dataset~\cite{wang2015accurate} & 8 & 199 & OPLS2.1 \\
FEP+ Dataset~\cite{harder2016opls3} & 8 & 199 & OPLS3 \\
FEP+ Dataset~\cite{roos2019opls3e} & 8 & 199 & OPLS3e \\
FEP+ Dataset~\cite{song2019using} & 8 & 199 & AMBER18 \\
Fragment Optimization~\cite{steinbrecher2015accurate} & 8 & 96 & OPLS2.1 \\
Scaffold Hopping~\cite{wang2017accurate} & 6 & 21 & OPLS3 \\
Macrocycles~\cite{yu2017accurate} & 7 & 33 & OPLS3 \\
\hline
\multicolumn{4}{|c|}{Further Suggested Datasets} \\
\hline
Cucurbit[7]uril (CB7)~\cite{mobley2017predicting} & 1 & 15 & NA \\
Deep cavity cavitand~\cite{mobley2017predicting} & 2 & 19 & NA \\
T4 Lysozyme~\cite{mobley2017predicting} & 2 & 20 & NA \\
Merck benchmarking set~\cite{MCompChem2019Sep} & 5 & 169 & OPSL3 \\

\hline
\label{tab:benchmarks}
\end{tabular}

>>>>

\paragraph{Several choices of restraints are possible.}
\todo[inline, color={blue!20}]{@author: Check references}
In practice, a variety of types of restraints are common, from simple harmonic distance restraints between the ligand and the protein~\cite{mobley2006use}, to flat-bottom restraints which work similarly but only exert a force if the ligand leaves a specific region~\cite{}.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{paper/figures/fig10_equib_detection/fig10_equib.pdf}
    \caption{{\bf Automatic partitioning into equilibration and production regions.}
    \emph{Top:} The average (black line) standard deviation (shaded region) of the reduced potential $u^*$ over many independent replicate simulations started from the same initial conditions show a significant initial transient before relaxing to the true average potential energy.
    \emph{Middle:} A cumulative average (red) of the entire simulation data demonstrates a bias in the estimate of the average potential $<u>$ that persists far beyond the initial transient relaxation time, while averaging a production region that omits the initial transient (blue) eliminates this bias.
    \emph{Bottom:} A simple automated approach to determination of the equilibration boundary $t_0$ suggested in~\cite{chodera2016simple} examines the number of effectively uncorrelated samples $N_{\mathrm{eff}_{[t_0,T]}} = (T-t_0) / g_{[t_0,T]}$ within the proposed production region $[t_0,T]$ by computing the statistical inefficiency $g_{[t_0,T]}$ for that simulation interval.
    $N_{\mathrm{eff}_{[t_0,T]}}$ has the general behavior of increasing linearly as $t_0$ moves backwards from $T$ until the initial transient relaxation is encountered, at which point it plunges.
    The optimal equilibration boundary $t_0$ is selected to maximize the number of uncorrelated samples.
    \emph{Figure adapted from~\cite{chodera2016simple}.}
    \todo[inline]{JDC will update this figure with something similar for a real alchemical free energy calculation.}
    }
    \label{fig:automatic-equilibration-detection}
\end{figure}

@@

A simple approach to automatically partitioning simulation data into equilibration and production regions is described in~\cite{chodera2016simple} (illustrated in Figure~\ref{fig:automatic-equilibration-detection}).
Suppose we have a simulation of length $T$ consisting of correlated data.
Here, the goal of the post-processing step is to select the equilibration boundary $t_0 \in [0, T]$ so as to \emph{maximize} the number of effectively uncorrelated samples remaining in the production region $N_{[t_0,T]}$, which is defined as
\begin{eqnarray}
N_{[t_0,T]} &=& \frac{T - t_0}{g_{[t_0,T]}}
\end{eqnarray}
where $g_{[t_0,T]}$ is the \emph{statistical inefficiency} of a timeseries $a_t$, described in more detail below.
Conveniently, this procedure also produces the information necessary to decorrelate the simulation data for estimating the free energy differences, a requisite next step in analysis.
This approach is implemented within the MBAR~\cite{kylebeauchamp2019choderalab} and alchemlyb~\cite{daviddotson2019alchemistry} packages, and is highly recommended for standard practice.

@@

Robust estimation of $C_t$ for $t \sim T$ is difficult due to growth in statistical error, so common estimators of $g$ make use of several additional properties of $C_t$ to provide useful estimates (see \emph{Practical Computation of Statistical Inefficiencies} in~\cite{chodera2016simple} for a detailed discussion).
\todo[inline]{JDC: Describe recommended statistical inefficiency computation algorithm.}
We recommend using the robust statistical inefficinecy computation routines available within the MBAR~\cite{kylebeauchamp2019choderalab} and alchemlyb~\cite{daviddotson2019alchemistry} packages.

\paragraph{Subsampling data to generate uncorrelated samples}
Once the statistical inefficiency $g$ has been estimated, it is straightforward to subsample the correlated timeseries simulation data to produce effectively uncorrelated data that can ba fed to estimators of the free energy.
Suppose the correlated timeseries is $\{a_t\}_{t=1}^T$; we can form a new timeseries of $N_{\mathrm{eff}} \approx T / g$ effectively uncorrelated samples by selecting a subset of indices $\{ \: t = \mathrm{round}((n-1) \, g) \: | \: n \in \mathrm{range}(1,\ldots ,N) \: \}$ where $\mathrm{round(x)}$ denotes rounding to the nearest integer.

@@

\paragraph{Cautions and considerations}
Reliable estimation of the statistical inefficiency is difficult, and estimates will not generally be as precise (in a relative error sense) as averages.
To ensure there is sufficient data available for reliable decorrelation and estimation of free energy differences, it is recommended that the effective number of uncorrelated samples $N_{\mathrm{eff}} \ge \approx 50$.

@@

The simplest method for calculating free energy
differences from simulations is the so-called \textit{Zwanzig
relationship}~\cite{zwanzig1954hightemperature}, also called one-sided exponential reweighting (EXP), or simply free energy perturbation, thought that term later grew to encompass all ways of calculating free energy differences.

The free energy between two states defined by two different potential energy functions 
$U_0(\vec{q})$ and $U_1(\vec{q})$ over coordinate space $\vec{q}$ can be calculated as:
\begin{eqnarray}
\Delta f_{01} & = & \ln \expect{e^{-(U_1(\vec{q}) - U_0(\vec{q}))}}_0 =  \ln \expect{e^{-\Delta U(\vec{q})} }_0
\end{eqnarray}\label{eqn.zwanzig}
In words, we take the samples generated during our run with the potential energy function $U_0(\vec{q})$, recalculate what the energy difference would be to switch to potential energy function $U_1(\vec{q})$, and exponentially average them to get the free energy.  Here, the exponential average consists in multiplying by -1, exponentiating, averaging, taking the logarithm, and multiplying by -1 again.  This almost requires no extra code functionality to perform; one need only to save a full precision trajectory, and run an unmodified molecular simulation code using the $U_1$ in order to calculate the new energies.  The analysis can be written in a line of code.  We note that this method is even more general, in that the instantaneous work to change the potential energy function from $U_0$ to $U_1$ can be replaced by the non-reversible work $W$ to make the same change some conditions~\cite{xx} 
Although the Zwanzig equation is formally correct (as long as the two states considered sample the same phase space volume), it has some very important numerical issues that mean that it is usually when performing a standard free energy calculation for a small molecule.~\cite{shirts2005comparison,lu2003appropriate}  One can show that if the standard deviation of the difference $\Delta U(\vec{q}) = U_1(vec{q})-U_2(\vec{q})$ over all sampled $\vec{q}$ is large (which in this case, means only several times $k_BT)$, then very few samples contribute to the average, and the answer will be both biased and extremely noisy~\cite{xx}.  There are a number of applications where it is useful, but those usually involve investigating the dependence of the free energy on small changes in parameters, not changing molecules or even single atoms in molecules~\cite{xx}

@@

Many studies have demonstrated both the
theoretical and practical superiority of BAR over EXP in molecular
simulations~\cite{shirts2005comparison,lu2003appropriate}, and BAR converges to EXP in the limit that all samples are from a
single state~\cite{bennett1976efficient,shirts.bennett}. Significantly less
overlap between the configurational space of each state is required to
converge than for EXP, though some overlap must still exist.

@@

Various numerical integration schemes are possible, but the trapezoid
rule provides a simple and robust scheme.  All types of
numerical integration can be written as:
\[ \Delta f \approx \sum_{k=1}^{K} w_k
\expect{\frac{du(\lambda,\vec{q})}{d\lambda}}_{k} \] where the weights
$w_k$ correspond to a particular choice of numerical integration.
Researchers have tried a large number of different integration
schemes~\cite{resat1993studies,jorge2010effect,shyu2009reducing}. For However, many other integration choices require specific choices of $\lambda$
to minimize bias, which makes them unsuitable when the intermediates
have widely-varying levels of uncertainty. For example, integrating a cubic spline interpolation provided negligble benefits~\cite{paliwal2011benchmark}. For starting researchers, we therefore recommend a simple trapezoid rule scheme, as it allows for maximal flexibility in which values of $\lambda$ are simulated.  Because fitting to higher
order polynomials can have numerical instabilities for some energy functions, and because alternate functional forms might only be appropriate
with some types of transformations, expertise and experience is
required to perform such numerical integration modifications.  In practice, adding 2-3 more intermediates is sufficient to match the performance of these more complicated numerical quadrature schemes.

@@

MBAR is provably the lowest variance asymptotically unbiases estimator of the free energies given the energies of the samples~\cite{zqtan}. It also provides an uncertainty estimate, which has been shown to be accurate as long as there are sufficient samples at each state~\cite{paliwal2011benchmark}.

MBAR can also be thought of as the Zwanzig estimator of the free energy to state $i$ where the sampled distribution is the \textit{mixture distribution} of all the other samples thrown together in one "pot", defined as $p_m(\vec{q} = N^{-1} \sum_k  N_k \exp(f_k-u_k\vec{q})$, the weighted average of all the individual normalized probability distributions that are performed.~\cite{shirts2003equilibrium}.

@@

\begin{itemize}
\item MBAR recommended if all energy differences are available. It is provably the lowest variance free energy estimate given samples from multiple states.
\item BAR is essentially just as good as MBAR for highly optimized $\lambda$ intermediates.  Specifically, if the $\lambda$s are chosen such that intermediate states have moderate overlap with their neighbors (i.e. between $i$ and $i+1$ and between $i$ and $i-1$, they will \textit{not} have significant overlap with their next nearest neighbors $i+2$ and $i-2$. Thus MBAR doesn't actually get significant information from these energy differences, so one might as well not even calculate them, and just perform BAR between nearest neighbors.~\cite{paliwal2011benchmark} 
\item TI usually gives similar values as MBAR implemented with sufficient numbers of intermediates, but quadrature error hard to estimate beforehand.~\cite{paliwal2011benchmark}
\item WHAM is an approximation to MBAR, and there are no compelling reasons it should be used. 
\item Other variants useful in special circumstances (e.g. Z. Tan stochastic version), but beyond the scope of a Best Practices article.
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%
%              Terminology                  %
%%%%%%%%%%%%%%%%%%%%
\section{Terminology and abbreviations}
\label{sec:tem-abbrev}
\begin{itemize}
\item Feature Box covering major technical terms and abbreviations
\item Examples:
\begin{itemize}
\item EXP, BAR, MBAR
\item Double decoupling, single-topology, dual-topology, hybrid-topology, coupled-topology
\item FEP (free energy peturbation), alchemical, AFE (alchemical free energy)
\end{itemize}
\end{itemize}
%%%%%%%%%%%%%%%%%%%%
%                Software                    %
%%%%%%%%%%%%%%%%%%%%
\section{Available software -- a summary}
\todo[inline]{Samar: write section}
\label{sec:software}
There are several softwares availble for carrying out alchemical free energy calculations. In this section we discuss some of the popular commercial and free tools along with their features. 
\begin{itemize}
\item Commercial:
   \begin{itemize}
    \item FEP+ is a tool offered by Schrodinger Inc. under a commercial license. It has an intuitive GUI which makes it easier for non-experts to run alchemical free energy calculations and analyze the results. It runs DESMOND MD package under the hood and hence parallelizes very well on the GPUs. 
    \end{itemize}
\item Free or low-cost for academics / commercial for industry:
	\begin{itemize}
	\item CHARMM has a variety of tools developed over the years. PERT module can be used to define initial and final states and define the intermediate lambda points. FREN and BAR modules can be used to analyze the data after the MD run. Lambda-dynamics based free energy calulation can be carried out using the BLOCK module.  
	\item TIES and AMBER FEW? (Peter Coveney)
	\item AMBER, including its new pmemd.cuda version support free energy calculation. 
	\end{itemize}
\item Free (libre) open source:
	\begin{itemize}
	\item PLUMED is an open source tool which enables the usage of a variety of MD engines. It is designed as a plugin for MD packages such that it analyzes the trajectory on the fly. It also offers a VMD based plugin for the computation of collective variables.   	
	\item SIRE  is a free, open source, multiscale molecular simulation framework, written to allow computational modellers to quickly prototype and develop new algorithms for molecular simulation and molecular design. 
	\item YANK is a tool developed by John Chodera and group on the top of OpenMM MD package. It allows the users to write their inputs in easy-to-use YAML format.
	\item GROMACS is a molecular simulation package with a significant number of free energy methods implementations. The LiveCOMS GROMACS tutorial has an example free energy calculation~\cite{Lemkul2018From}.
	\item pmx for mutations
	\end{itemize}
\item Setup tools
	\begin{itemize}
	\item FESetup: AMBER, gromacs, Sire
	\item Lomap/Lomap2 : Relative alchemical transformation graph planning
	\item CHARMM-GUI is a web based tool for setting up a variety of MD simulations. It can be used to generate CHARMM sctipts for solvation and ligand-binding free energy calculations.
	\end{itemize}
\item Analysis tools:
	\begin{itemize}
	\item Free Energy Workflows: Sire-specific free energy map analysis using weighted path averages
	\url{https://github.com/michellab/freenrgworkflows}
	\item Alchemlyb: Multipackage free energy analysis
	\url{https://github.com/alchemistry/alchemlyb}
	\item pymbar: MBAR implementation, but have to roll your own analysis wrapper
	\url{https://github.com/choderalab/pymbar}
	\end{itemize}
\end{itemize}

@@

\begin{table}
\caption{Overview of benchmark datasets}
\begin{tabular}{cccc}

\textbf{Publication} & \textbf{Targets} & \textbf{Ligands} & \textbf{Force Field} \\
\hline
\multicolumn{4}{|c|}{D3R Grand Challenges~\cite{D3R}} \\
\hline
GC3~\cite{gaieb2019d3r} & 6 & 266 & various \\
GC2~\cite{gaieb2018d3r} & 1 & 102 & various \\
GC2015~\cite{gathiaka2016d3r} & 2 & 215 & various \\
\hline
\multicolumn{4}{|c|}{SAMPL Challenges~\cite{SAMPL}} \\
\hline
SAMPL6~\cite{rizzi2018overview} & 3 & 21 & various \\
SAMPL5~\cite{yin2017overview} & 3 & 22 & various \\
SAMPL4~\cite{muddana2014sampl4} & 2 & 23 & various \\
\hline
\multicolumn{4}{|c|}{Schrödinger Datasets} \\
\hline
FEP+ Dataset~\cite{wang2015accurate} & 8 & 199 & OPLS2.1 \\
FEP+ Dataset~\cite{harder2016opls3} & 8 & 199 & OPLS3 \\
FEP+ Dataset~\cite{roos2019opls3e} & 8 & 199 & OPLS3e \\
FEP+ Dataset~\cite{song2019using} & 8 & 199 & AMBER18 \\
Fragment Optimization~\cite{steinbrecher2015accurate} & 8 & 96 & OPLS2.1 \\
Scaffold Hopping~\cite{wang2017accurate} & 6 & 21 & OPLS3 \\
Macrocycles~\cite{yu2017accurate} & 7 & 33 & OPLS3 \\
\hline
\multicolumn{4}{|c|}{Further Suggested Datasets} \\
\hline
Cucurbit[7]uril (CB7)~\cite{mobley2017predicting} & 1 & 15 & NA \\
Deep cavity cavitand~\cite{mobley2017predicting} & 2 & 19 & NA \\
T4 Lysozyme~\cite{mobley2017predicting} & 2 & 20 & NA \\
Merck benchmarking set~\cite{MCompChem2019Sep} & 5 & 169 & OPSL3 \\

\hline
\label{tab:benchmarks}
\end{tabular}
\end{table}
